# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
if(sum(list.files()=="activity.csv") == 0){
  fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
  activity <- unz(fileURL,fileName="activity.csv",open="r")
}
activity <- read.csv("activity.csv")

activity$date = as.Date(activity$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

```{r}
steps_per_day <- aggregate(activity$steps, by=list(activity$date), sum)
names(steps_per_day)[1] <- "date"
names(steps_per_day)[2] <- "steps"
mean_steps <- mean(steps_per_day$steps,na.rm=TRUE)
median_steps <- median(steps_per_day$steps,na.rm=TRUE)
hist(steps_per_day$steps)
```

The mean total number of steps taken per day is **`r as.integer(mean_steps)`**.  
The median total number of steps taken per day is **`r median_steps`**.  

## What is the average daily activity pattern?
```{r}
steps_per_interval <- aggregate(activity$steps, by=list(activity$interval), mean, na.rm=TRUE)
names(steps_per_interval)[1] <- "interval"
names(steps_per_interval)[2] <- "steps"
plot(steps_per_interval$interval,steps_per_interval$steps,type="l")
max_interval <- steps_per_interval$interval[which.max(steps_per_interval$steps)]
```

The five minute interval that contains the highest number of stps is interval **`r max_interval`**.  

## Imputing missing values
```{r}
na_data <- apply(is.na(activity),1,sum)
na_rows <- length(na_data) - length(na_data[na_data==0])
```

The total number of missing values in the dataset are **`r na_rows`**.  

```{r}
activity$steps_nona = activity$steps
for (i in 1:length(activity$steps)){
  if (is.na(activity$steps[i])) {
    activity$steps_nona[i] = mean(activity$steps, na.rm = TRUE)
  }
}

steps_per_day_nona <- aggregate(activity$steps_nona, by=list(activity$date), sum)
names(steps_per_day_nona)[1] <- "date"
names(steps_per_day_nona)[2] <- "steps"
mean_steps_nona <- mean(steps_per_day_nona$steps)
median_steps_nona <- median(steps_per_day_nona$steps)
hist(steps_per_day_nona$steps)
```

The mean total number of steps taken per day is **`r as.integer(mean_steps_nona)`**.  
The median total number of steps taken per day is **`r as.integer(median_steps_nona)`**.  
There is little difference between the calculation with filled-in missing values and the calculation ignoring missing values.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
activity$day <- weekdays(activity$date)
weekdays_list <- c("Monday","Tuesday","Wednesday","Thursday","Friday")
activity$daytype <- lapply(activity$day, function(x) ifelse((x == "Saturday" || x == "Sunday"),x<-"Weekend",x<-"Weekday"))
```

```{r}
steps_per_interval_daytype.wd <- aggregate(steps ~ interval, data = activity, subset = activity$daytype == 
        "Weekday", FUN = mean)
steps_per_interval_daytype.we <- aggregate(steps ~ interval, data = activity, subset = activity$daytype == 
        "Weekend", FUN = mean)
par(mfrow=c(2,1))
plot(steps_per_interval_daytype.wd$interval,steps_per_interval_daytype.wd$steps,type="l")
plot(steps_per_interval_daytype.we$interval,steps_per_interval_daytype.we$steps,type="l")



```

The main difference between weekend and weekday activity is that weekday activity begins earlier in the day.